Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-06 manuscript/supplemental_material.Rnw
remove empty table
huge gene table now has more information
include annotation
include links
include human alignment where available

change link color to black

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=reading,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","palcenta_go_results_mf")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf"
                              ))) {
    load(a)
}
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
@

@@

\begin{table}
  \begin{tabular}
  \end{tabular}
\end{table}

@@

<<genes.expression,echo=FALSE,results="hide">>=
genes.expression.table <- function(species.name,min.fpkm=50) {
    temp <- 
        combined.fpkm[species==species.name][order(-mean_fpkm)][,list(name_or_id,
                          human_name,mean_fpkm)]
    temp <- temp[!duplicated(name_or_id)][mean_fpkm>=min.fpkm]
    setnames(temp,c("Gene name","Human Name","FPKM"))
    if (species.name=="homo sapiens") {
        temp <- temp[,-2,with=FALSE]
    }
    return(temp)
}
@

\section{Gene Expression}
<<top.expressed.genes,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    this.caption <- paste0("Expression of all genes with FPKM $\\ge 50$ in \\textit{",
                           capfirst(this.species),"}")
    cat(paste0("\\subsection{",this.caption,"}\n\n"))
    print(xtable(genes.expression.table(species.name=this.species),
                 caption=this.caption,
                 label=paste0("tab:all_",gsub(" ","_",this.species),"_placenta")
                 ),
          booktabs=TRUE,
          tabular.environment="longtable",
          floating=FALSE,
          caption.placement="top",
          size="\\tiny"
          )
}
@

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=reading,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\newenvironment{narrow}[2]{%
  \begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\leftmargin}{#1}%
      \setlength{\rightmargin}{#2}%
      \setlength{\listparindent}{\parindent}%
      \setlength{\itemindent}{\parindent}%
      \setlength{\parsep}{\parskip}}%
  \item[]}{\end{list}}
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset","placenta_core_transcriptome","placenta_go_results_bp","palcenta_go_results_mf","hugo_nomenclature","gene_annotation")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees",
                              "housekeeping_genes_superset",
                              "placenta_core_transcriptome",
                              "placenta_go_results_bp",
                              "placenta_go_results_mf",
                              "hugo_nomenclature",
                              "gene_annotation"
                              ))) {
    load(a)
}
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
setkey(hugo.nomenclature,"Approved Symbol")
@

@@

@@

<<genes.expression,echo=FALSE,results="hide">>=
genes.expression.table <- function(species.name,min.fpkm=50) {
    alignment.species <- species.name
    if (alignment.species=="pan paniscus") {
        alignment.species <- "pan troglodytes"
    } else if (alignment.species=="ateles fusciceps") {
        alignment.species <- "homo sapiens"
    } else if (grepl("spalax",alignment.species)) {
        alignment.species <- "mus musculus"
    }

    temp <- 
        combined.fpkm[species==species.name
                      ][order(-mean_fpkm)
                        ][,list(name_or_id,
                                human_alignment_symbol,
                                human_name,mean_fpkm)]
    temp <- temp[!duplicated(name_or_id)][mean_fpkm>=min.fpkm]
    temp[,used_alignment:=FALSE]
    temp[is.na(human_name) & !is.na(human_alignment_symbol),
         used_alignment:=TRUE]
    temp[is.na(human_name),human_name:=human_alignment_symbol]
    setkey(temp,"human_name")
    temp <- 
        hugo.nomenclature[temp][order(-mean_fpkm),
                                list(name_or_id,used_alignment,`Approved Symbol`,`Approved Name`,mean_fpkm)]
    temp[!is.na(name_or_id),
         name_or_id:=paste0("\\href{http://www.ensembl.org/",
                            capfirst(gsub(" ","_",alignment.species)),
                            "/Gene/Summary?g=",name_or_id,"}{",
                            xtable.sanitize.text(name_or_id),"}")]
    temp[!is.na(`Approved Name`),
         `Approved Name`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Name`),"}")]
    temp[!is.na(`Approved Symbol`),
         `Approved Symbol`:=paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",`Approved Symbol`,"}{",xtable.sanitize.text(`Approved Symbol`),"}")]
    temp[used_alignment==TRUE,`Approved Symbol`:=paste0("(",`Approved Symbol`,")")]
    temp[,used_alignment:=NULL]
    setnames(temp,c("Gene symbol","Human Symbol","Human Name","FPKM"))
    if (species.name=="homo sapiens" || species.name=="ateles fusciceps") {
        temp[,`Human Symbol`:=NULL]
    }
    return(temp)
}
@

\section{Gene Expression}
<<top.expressed.genes,echo=FALSE,results="asis">>=
for (this.species in names(species.names)) {
    this.caption <- paste0("\\textit{",
                           capfirst(this.species),"} Placental Gene Expression of all genes with FPKM $\\ge 50$")
    cat(paste0("\\subsection{",this.caption,"}\n\n"))
    print(xtable(genes.expression.table(species.name=this.species),
                 caption=this.caption,
                 label=paste0("tab:all_",gsub(" ","_",this.species),"_placenta")
                 ),
          booktabs=TRUE,
          tabular.environment="longtable",
          floating=FALSE,
          caption.placement="top",
          sanitize.text.function=genes.table.sanitize,
          size="\\tiny"
          )
}
@
